---
title: "Findings"
output: 
  html_document:
    toc: true
    toc_float: true
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(dplyr)
library(ggplot2)
library(plotly)
library(shiny)
source("process_data.R")

```

\

This Findings page presents the key findings from the regression analysis of global suicide rates. It begins by comparing three models with increasing complexity to evaluate their performance. The best-performing model is then examined in more detail, followed by a discussion of how the effects of key predictors vary across world regions.

\

## Model Selection

To explore how key predictors influence suicide rates, three linear regression models were fitted using different specifications of the response variable and predictors. Below are their formulas in R code:

\

### Model 0

```{r model_0}
model_0 <- lm(SuicideRate ~ 
                GDP_per_Capita + GINI_Index + HealthSpendingPercent, 
              data = data_merged)

```

Model 0 is a baseline model that uses the untransformed suicide rate as the response and includes the three main numerical predictors.

\

### Model 1

```{r model_1}
model_1 <- lm(SuicideRate_sqrt ~ 
                GDP_per_Capita + GINI_Index + HealthSpendingPercent, 
              data = data_merged)

```

Model 1 applies a square root transformation to the response variable to improve linearity and variance stability.

\

### Model 2

```{r model_2}
model_2 <- lm(SuicideRate_sqrt ~ 
                (GDP_per_Capita + GINI_Index + HealthSpendingPercent)*Region, 
              data = data_merged)

```

Model 2 adds the `Region` variable and interaction terms to capture regional differences in the effects of predictors.

\

### Performance Comparison

<table style="width:100%; border-collapse:collapse; margin-bottom: 1em;">
  <thead>
    <tr>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Model</th>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Response</th>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Predictors</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">RSS</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">R-squared</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">Adj. R-squared</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">F-test p-value</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>0</td>
      <td>SuicideRate</td>
      <td>GDP, Gini, Health</td>
      <td style="text-align:right;">88435.973</td>
      <td style="text-align:right;">0.136</td>
      <td style="text-align:right;">0.134</td>
      <td style="text-align:right;">0.000</td>
    </tr>
    <tr>
      <td>1</td>
      <td>√SuicideRate</td>
      <td>GDP, Gini, Health</td>
      <td style="text-align:right;">1699.234</td>
      <td style="text-align:right;">0.175</td>
      <td style="text-align:right;">0.174</td>
      <td style="text-align:right;">0.000</td>
    </tr>
    <tr style="background-color: #f0f0f0;">
      <td>2</td>
      <td>√SuicideRate</td>
      <td>GDP, Gini, Health + Region + Interactions</td>
      <td style="text-align:right;">1278.190</td>
      <td style="text-align:right;">0.380</td>
      <td style="text-align:right;">0.370</td>
      <td style="text-align:right;">0.000</td>
    </tr>
  </tbody>
</table>

**Table 1:** Coefficient summary table for Model 0, Model 1, and Model 2. The table presents coefficient estimates, standard errors, t-values, and associated p-values for each term in each model.

Note that Model 2 shows the best performance, with the lowest residual sum of squares and highest $R^2$ and adjusted $R^2$ values, justifying the inclusion of regional effects and interactions.</p>

\

## Final Model Summary

### Coefficient Estimates

<table style="width:100%; border-collapse:collapse; margin-bottom: 1em;">
  <thead>
    <tr>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Term</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">Estimate</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">Std. Error</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">t value</th>
      <th style="text-align:right; border-bottom:2px solid #ccc;">t-Test p-value</th>
    </tr>
  </thead>
  <tbody>
    <tr><td>Intercept</td><td style="text-align:right;">0.377</td><td style="text-align:right;">0.411</td><td style="text-align:right;">0.917</td><td style="text-align:right;">0.359</td></tr>
    <tr><td>GDP per Capita</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">1.591</td><td style="text-align:right;">0.112</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Gini Index</td><td style="text-align:right;">4.551</td><td style="text-align:right;">0.966</td><td style="text-align:right;">4.711</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Health Spending</td><td style="text-align:right;">0.048</td><td style="text-align:right;">0.038</td><td style="text-align:right;">1.265</td><td style="text-align:right;">0.206</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Region Americas</td><td style="text-align:right;">2.558</td><td style="text-align:right;">0.703</td><td style="text-align:right;">3.639</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Region Eastern Mediterranean</td><td style="text-align:right;">1.158</td><td style="text-align:right;">1.070</td><td style="text-align:right;">1.082</td><td style="text-align:right;">0.279</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Region Europe</td><td style="text-align:right;">5.938</td><td style="text-align:right;">0.502</td><td style="text-align:right;">11.836</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Region South-East Asia</td><td style="text-align:right;">-1.583</td><td style="text-align:right;">1.161</td><td style="text-align:right;">-1.364</td><td style="text-align:right;">0.173</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Region Western Pacific</td><td style="text-align:right;">5.499</td><td style="text-align:right;">1.081</td><td style="text-align:right;">5.087</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>GDP per Capita:Region Americas</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">-1.021</td><td style="text-align:right;">0.308</td></tr>
    <tr><td>GDP per Capita:Region Eastern Mediterranean</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">-1.353</td><td style="text-align:right;">0.176</td></tr>
    <tr><td>GDP per Capita:Region Europe</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">-1.558</td><td style="text-align:right;">0.119</td></tr>
    <tr><td>GDP per Capita:Region South-East Asia</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">-0.410</td><td style="text-align:right;">0.682</td></tr>
    <tr><td>GDP per Capita:Region Western Pacific</td><td style="text-align:right;">0.000</td><td style="text-align:right;">0.000</td><td style="text-align:right;">-0.659</td><td style="text-align:right;">0.510</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Gini Index:Region Americas</td><td style="text-align:right;">-7.475</td><td style="text-align:right;">1.431</td><td style="text-align:right;">-5.222</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Gini Index:Region Eastern Mediterranean</td><td style="text-align:right;">-1.385</td><td style="text-align:right;">2.968</td><td style="text-align:right;">-0.467</td><td style="text-align:right;">0.641</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Gini Index:Region Europe</td><td style="text-align:right;">-10.736</td><td style="text-align:right;">1.208</td><td style="text-align:right;">-8.891</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Gini Index:Region South-East Asia</td><td style="text-align:right;">5.732</td><td style="text-align:right;">3.205</td><td style="text-align:right;">1.789</td><td style="text-align:right;">0.074</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Gini Index:Region Western Pacific</td><td style="text-align:right;">-14.362</td><td style="text-align:right;">2.629</td><td style="text-align:right;">-5.463</td><td style="text-align:right;">0.000</td></tr>
    <tr><td>Health Spending:Region Americas</td><td style="text-align:right;">0.066</td><td style="text-align:right;">0.048</td><td style="text-align:right;">1.363</td><td style="text-align:right;">0.173</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Health Spending:Region Eastern Mediterranean</td><td style="text-align:right;">-0.215</td><td style="text-align:right;">0.060</td><td style="text-align:right;">-3.601</td><td style="text-align:right;">0.000</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Health Spending:Region Europe</td><td style="text-align:right;">-0.141</td><td style="text-align:right;">0.041</td><td style="text-align:right;">-3.417</td><td style="text-align:right;">0.001</td></tr>
    <tr><td>Health Spending:Region South-East Asia</td><td style="text-align:right;">-0.051</td><td style="text-align:right;">0.079</td><td style="text-align:right;">-0.643</td><td style="text-align:right;">0.521</td></tr>
    <tr style="background-color:#f0f0f0;"><td>Health Spending:Region Western Pacific</td><td style="text-align:right;">0.110</td><td style="text-align:right;">0.056</td><td style="text-align:right;">1.974</td><td style="text-align:right;">0.049</td></tr>
  </tbody>
</table>

**Table 2:** Coefficient summary table for Model 2, which models square root-transformed suicide rates using GDP per capita, Gini index, health spending, region, and interaction terms between region and numerical predictors. The table presents coefficient estimates, standard errors, t-values, and associated p-values for each term in the model.

Among the main effects in Model 2, Gini index and region indicators for Americas, Europe, and Western Pacific showed significant associations with the square root of suicide rates (p-value < 0.05). More importantly, multiple interaction terms were also significant, including the interaction between Gini index and Americas, Europe, and Western Pacific, as well as the interactions between health spending and Eastern Mediterranean, Europe, and Western Pacific (p-value < 0.05). These results indicated that the effects of Gini index and health spending on suicide rates were not constant across regions, but rather varied significantly depending on the geographic context.

\

### Regional Variations

<table style="width:100%; border-collapse:collapse; margin-bottom: 1em;">
  <thead>
    <tr>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Region</th>
      <th style="text-align:left; border-bottom:2px solid #ccc;">GDP per Capita</th>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Gini Index</th>
      <th style="text-align:left; border-bottom:2px solid #ccc;">Health Spending</th>
    </tr>
  </thead>
  <tbody>
    <tr>
      <td>Africa</td>
      <td>0.000</td>
      <td style="background-color:#f0f0f0;">4.551</td>
      <td>0.048</td>
    </tr>
    <tr>
      <td>Americas</td>
      <td>0.000</td>
      <td style="background-color:#f0f0f0;">-2.924</td>
      <td>0.114</td>
    </tr>
    <tr>
      <td>Eastern Mediterranean</td>
      <td>0.000</td>
      <td>3.166</td>
      <td style="background-color:#f0f0f0;">-0.167</td>
    </tr>
    <tr>
      <td>Europe</td>
      <td>0.000</td>
      <td style="background-color:#f0f0f0;">-6.185</td>
      <td style="background-color:#f0f0f0;">-0.093</td>
    </tr>
    <tr>
      <td>South-East Asia</td>
      <td>0.000</td>
      <td>10.283</td>
      <td>-0.003</td>
    </tr>
    <tr>
      <td>Western Pacific</td>
      <td>0.000</td>
      <td style="background-color:#f0f0f0;">-9.811</td>
      <td style="background-color:#f0f0f0;">0.158</td>
    </tr>
  </tbody>
</table>

**Table 3:** Summary table of the net effect of each numerical predictor on the square root of suicide rate across world regions. Highlighted cells indicate statistically significant effects (p-value < 0.05) based on Model 2's coefficient estimates and interaction terms.

As shown in Table 3, a one unit increase in Gini index was associated with a 4.551 increase in the square root of suicide rate in Africa. However, this effect varied significantly across regions. In the Americas, the net effect was -2.924, suggesting a decrease in suicide rates with higher inequality. In Europe, the net effect was -6.185, indicating a more substantial negative association. The Western Pacific showed the most extreme net effect of -9.811, further highlighting how higher Gini index corresponded to markedly lower suicide rates in this region. For health spending, the main effect was 0.048 in Africa. In the Western Pacific, the interaction term was 0.110, leading to a net effect of 0.158, suggesting that a one percentage increase in health spending corresponded to approximately 0.158 unit increase in the square root of suicide rate. In contrast, the net effect in the Eastern Mediterranean was -0.167, and in Europe it was -0.093, both indicating negative associations between health spending and suicide rates. 
